Machine learning enhanced empirical potentials for metals and alloys

### 2020-05-20 doc/papers/ml-eam/manuscript.tex
Initial version for cpc.

\documentclass[prb,reprint,superscriptaddress]{revtex4-2}

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external
\raggedbottom           % don't add extra vertical space
\begin{comment}
\end{comment}

\bibliographystyle{apsrev4-2}

\begin{document}

@@

\author{Xin Chen}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Li-Fang Wang}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Xing-Yu Gao}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Ya-Fan Zhao}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{De-Ye Lin}
\email{lindeye0716@163.com}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Wei-Dong Chu}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Hai-Feng Song}
\email{song\_haifeng@iapcm.ac.cn}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

@@

\maketitle

@@

The Rose constraint incoporates the universal equation of state (Rose et al) 
into the total loss function. The Mishin-modified equation 
\begin{equation}
\label{eq:rose_eos}
E^\mathrm{rose}(x) = E_{0}\left[
    1 + \alpha x + \beta \alpha^3 x^3 \frac{2x + 3}{(x - 1)^2} \right]
    e^{-\alpha x}
\end{equation}
is used because the original form tends to underestimate energies under high 
pressures. In Equation \ref{eq:rose_eos}, $E_{0}$ is the energy of the 
equilibrium structure, $x = a / a_{0} - 1$ is the relative isotropic scaling 
factor ($a$ is the lattice constant), $\beta$ is a chosen parameter and 
\begin{equation}
\label{eq:rose_alpha}
\alpha = \sqrt{-\frac{9 V_{0} B }{E_{0}}}
\end{equation}
where $V_0$ is the equilibrium volume and $B$ is the corresponding bulk modulus. 
The adoption of the Rose constraint guarantees the exact predictions of bulk 
modulus and the energy-volume curve. The loss of the Rose EOS constraint 
$\mathbf{L}^{\mathrm{Rose}}$  is measured as the 2-norm of the energy 
differences between $E^{\mathrm{rose}}(x)$ and their corresponding predicted 
$E(x)$: 
\begin{equation}
\label{eq:rose_loss}
\mathbf{L}^{\mathrm{Rose}} = \sum_{\Omega}{
    \sqrt{\sum_{i}{\left(E(x_i) - E^{\mathrm{rose}}(x_i)\right)^2}}
}
\end{equation}
Here $\Omega$ denotes all selected crystals. In this work, for each included 
crystal, we use the same choices of $x$, which determines the fitting region: 
$x_{i} = x_{0} + \Delta x \cdot i, x_{0} = -0.1, N_{t}^{\mathrm{max}} = 20, 
\Delta x=0.01$. (-10\% to +10\% of the equilibrium volume). In this work, 
$\beta^{\mathrm{Ni}}$ is 0.005 and $\beta^{\mathrm{Mo}}$ is 0.008.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EAM/ADP - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.45]{tensoralloy.png}
\caption{\label{fig:tensoralloy} The architecture of TensorAlloy.}
\end{figure}

@@

For the second step, VAP also plays the key role. Only very few specific 
structures are needed to form physical constraints. Thus, the batching scheme is 
no longer a prerequiste. Instead, we use the codes of the 'prediction' phase of 
VAP for flexibility \cite{TensorAlloy}. For each type of physical constraint and 
each selected crystal structure, an independent computation graph shall be 
generated by reusing the model parameters.

To find optimal parameters of the funcitons in section \ref{sec:functions}, the 
overall loss function shall be defined. Just like other machine learning tasks, 
the mini-batch stochastic gradient descent algorithm is used to minimize the 
loss function. Equation \ref{eq:loss} demonstrates the loss function used in 
this work:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3\sum_{i}^{N_{b}}{N_i}}\sum_{i}^{N_b}{\sum_{j}^{N_i}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{rose}}\mathbf{L}^{\mathrm{Rose}} 
+ \chi_{\mathrm{elastic}}\mathbf{L}^{\mathrm{elastic}}
\end{align} 
where $N_{b}$ is the batch size and $N_i$ is the number of atoms in structure 
$i$. $\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and $\chi_{\mathrm{elastic}}$ 
are overall weights of force, rose EOS and elastic contributions. 
The Adam optimizer \cite{adam} is used to minimize Equation \ref{eq:loss}. 
In most cases, we use 0.01 as the initial learning rate and the batch size 
ranges from 20 to 50. Since there are very few adjustable parameters compared 
with neural networks, the optimization typically needs very few epochs to 
converge. When the optimization is finished, the corresponding Lammps setfl 
potential file will be exported.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EAM - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.57]{Ni_eam.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure*}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Alloy EOS - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.5]{alloy_eos.png}
\caption{\label{fig:alloy_eos} Energy-volume curves of Ni$_3$Mo and Ni$_4$Mo 
obtained with the original Zjw04 EAM, ML-EAM, ML-ADP, SNAP and DFT.}
\end{figure*}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% References
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\bibliography{manuscript.bib}

>>>>

%% This template can be used to write a paper for
%% Computer Physics Communications using LaTeX.
%% For authors who want to write a computer program description,
%% an example Program Summary is included that only has to be
%% completed and which will give the correct layout in the
%% preprint and the journal.
%% The `elsarticle' style is used and more information on this style
%% can be found at 
%% http://www.elsevier.com/wps/find/authorsview.authors/elsarticle.
%%
%%
%% \documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
\documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external
\usepackage{amssymb}

\newcounter{bla}
\newenvironment{refnummer}{%
\list{[\arabic{bla}]}%
{\usecounter{bla}%
 \setlength{\itemindent}{0pt}%
 \setlength{\topsep}{0pt}%
 \setlength{\itemsep}{0pt}%
 \setlength{\labelsep}{2pt}%
 \setlength{\listparindent}{0pt}%
 \settowidth{\labelwidth}{[9]}%
 \setlength{\leftmargin}{\labelwidth}%
 \addtolength{\leftmargin}{\labelsep}%
 \setlength{\rightmargin}{0pt}}}
{\endlist}

\journal{Computer Physics Communications}

\begin{document}
\begin{frontmatter}

@@

\author[a]{Xin Chen}
\author[a]{Li-Fang Wang}
\author[a]{Xing-Yu Gao}
\author[b,a]{Ya-Fan Zhao}
\author[b,a]{De-Ye Lin\corref{author}}
\author[a]{Wei-Dong Chu}
\author[a,b]{Hai-Feng Song\corref{author}}

\cortext[author] {Corresponding author.
\textit{E-mail address:} lindeye0716@163.com, song\_haifeng@iapcm.ac.cn}
\address[a]{Institute of Applied Physics and Computational Mathematics, 
Beijing 100088, China}
\address[b]{CAEP Software Center for High Performance Numerical Simulation, 
Beijing 100088, China}

@@

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Machine Learning \sep Alloy \sep Materials Modeling \sep Computational Physics 
\sep Empirical Potentials    
\end{keyword}

\end{frontmatter}

% Computer program descriptions should contain the following
% PROGRAM SUMMARY.

{\bf PROGRAM SUMMARY} \\
\begin{small}
\noindent
{\em Program Title: TensorAlloy} \\
{\em Licensing provisions(please choose one): LGPL}             \\
{\em Programming language: Python 3.7}                          \\
{\em Supplementary material:}                                   \\
{\em Journal reference of previous version: 
https://doi.org/10.1016/j.cpc.2019.107057}                      \\
{\em Does the new version supersede the previous version?:} YES \\
{\em Reasons for the new version: This new version is a significant extension to 
the previous version. Now machine learning approaches and physical constraints 
can be used together to tune empirical potentials (for example the embbeded atom 
method). Machine learning optimized empirical potentials can be almost as 
accurate as machine learning interaction potentials but run much faster} \\
{\em Summary of revisions:}*\\
{\em Nature of problem:} \\
  Optimizing empirical potentials with machine learning approaches and physical 
  constraints. \\
{\em Solution method:} \\
  The TensorAlloy program is built upon TensorFlow and the virtual-atom 
  approach. we successfully developed a route to express the embbeded atom 
  method and the angular-dependent potential within machine learning framework 
  in highly-vectorizable form and further enhanced the potentials with physical
  constraints. Machine learning can significantly boost their performances with 
  few training data.   \\
{\em Additional comments including Restrictions and Unusual features}\\
  This program needs TensorFlow 1.14.*. Neither newer or older TensorFlow is 
  supported. \\
\end{small}

@@

The Rose constraint incoporates the universal equation of state (Rose et al) 
into the total loss function. The Mishin-modified equation 
\begin{equation}
\label{eq:rose_eos}
E^\mathrm{rose}(x) = E_{0}\left[
    1 + \alpha x + \beta \alpha^3 x^3 \frac{2x + 3}{(x - 1)^2} \right]
    e^{-\alpha x}
\end{equation}
is used because the original form tends to underestimate energies under high 
pressures. In Equation \ref{eq:rose_eos}, $E_{0}$ is the energy of the 
equilibrium structure, $x = a / a_{0} - 1$ is the relative isotropic scaling 
factor ($a$ is the lattice constant), $\beta$ is a chosen parameter and 
\begin{equation}
\label{eq:rose_alpha}
\alpha = \sqrt{-\frac{9 V_{0} B }{E_{0}}}
\end{equation}
where $V_0$ is the equilibrium volume and $B$ is the corresponding bulk modulus. 
The adoption of the Rose constraint guarantees the exact predictions of bulk 
modulus and the energy-volume curve. The loss of the Rose EOS constraint 
$\mathbf{L}^{\mathrm{Rose}}$  is measured as the 2-norm of the energy 
differences between $E^{\mathrm{rose}}(x)$ and their corresponding predicted 
energy $E(x)$: 
\begin{equation}
\label{eq:rose_loss}
\mathbf{L}^{\mathrm{Rose}} = \sum_{\Omega}{
    \sqrt{\sum_{i}{\left(E(x_i) - E^{\mathrm{rose}}(x_i)\right)^2}}
}
\end{equation}
Here $\Omega$ denotes all selected crystals. In this work, for each included 
crystal, we use the same choices of $x$, which determines the fitting region: 
$x_{i} = x_{0} + \Delta x \cdot i, x_{0} = -0.1, N_{t}^{\mathrm{max}} = 20, 
\Delta x=0.01$. (-10\% to +10\% of the equilibrium volume). In this work, 
$\beta^{\mathrm{Ni}}$ is 0.005 and $\beta^{\mathrm{Mo}}$ is 0.008.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Arch
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.6]{tensoralloy.png}
\caption{\label{fig:tensoralloy} The architecture of TensorAlloy.}
\end{figure}

@@

For the second step, only very few specific 
structures are needed to form physical constraints. Thus, the batching scheme is 
no longer a pre-requiste. Instead, we use the codes of the 'prediction' phase of 
VAP for flexibility \cite{TensorAlloy}. For each type of physical constraint of 
each selected crystal structure, an independent computation graph shall be 
generated by reusing the model parameters. This can be easily achieved within 
our VAP framework. Hence, the complicated derivatives of the losses with respect 
to the model parameters can be solved by TensorFlow automatically.
For example, for the Ni-Mo system, if we select the elastic constants of $c_11$, 
$c_22$ and $c_44$ of fcc Ni and bcc Mo, and the Rose EOS of fcc Ni and bcc Mo as 
the physical constraints, thereshould be $2 \times 2$ separate computation 
graphs (\textrm{Elastic/fcc/Ni, Elastic/bcc/Mo, Rose/fcc/Ni, Rose/bcc/Mo}) under 
the scope 'Constraint'.

The total loss is just the sum of the two steps introduced above:
Equation \ref{eq:loss} demonstrates the loss function used in 
this work:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3\sum_{i}^{N_{b}}{N_i}}\sum_{i}^{N_b}{\sum_{j}^{N_i}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{rose}}\mathbf{L}^{\mathrm{Rose}} 
+ \chi_{\mathrm{elastic}}\mathbf{L}^{\mathrm{elastic}}
\end{align} 
where $N_{b}$ is the batch size and $N_i$ is the number of atoms in structure 
$i$. $\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and $\chi_{\mathrm{elastic}}$ 
are overall weights of force, rose EOS and elastic contributions.

To find optimal parameters of the funcitons in section \ref{sec:functions}, the 
overall loss function shall be defined. Just like other machine learning tasks, 
the mini-batch stochastic gradient descent algorithm is used to minimize the 
loss function. 
The Adam optimizer \cite{adam} is used to minimize Equation \ref{eq:loss}. 
In most cases, we use 0.01 as the initial learning rate and the batch size 
ranges from 20 to 50. Since there are very few adjustable parameters compared 
with neural networks, the optimization typically needs very few epochs to 
converge. When the optimization is finished, the corresponding Lammps setfl 
potential file will be exported.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EAM - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.4]{Ni_eam.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure*}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Alloy EOS - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.4]{alloy_eos.png}
\caption{\label{fig:alloy_eos} Energy-volume curves of Ni$_3$Mo and Ni$_4$Mo 
obtained with the original Zjw04 EAM, ML-EAM, ML-ADP, SNAP and DFT.}
\end{figure*}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% References
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\bibliographystyle{elsarticle-num}
\bibliography{manuscript.bib}
